### 2018-03-20 generativemodelsforscrna/outline.tex
updates for mtg

% !TEX program = pdflatex
\input{preamble/preamble.tex}
\input{preamble/preamble_math.tex}
\input{preamble/preamble_acronyms.tex}

% \linenumbers
\DeclareMathOperator*{\Categorical}{Categorical}

\title{Generative Models for scRNA}
\author{Jackson Loper}

\usepackage{amsthm}
\newtheorem{thm}{Theorem}
\newtheorem{conj}{Conjecture}

\begin{document}
\maketitle

\begin{description}
  \item[Status] Succesfully trained the new ``iterative model,'' evaluated it to some extent
  \item[Todo] Compare clusters against Allen, consider linear model, do more layers
\end{description}

\section{Overview}

The Allen Institute has developed a method for clustering, one that has apparently allowed them to distinguish 116 unique cell types using single cell RNA data.  The method follows the following iterative procedure:

\begin{algorithm}[H]
 \KwData{RNA expression for a collection of cells}
 \vspace{.1in}
  Run PCA on the data (e.g. take first 16 components)\;
  Run T-SNE on the resulting low-d representation\;
  Circle things that look like clusters\;
  \If{there is more than one cluster}{\For{each cluster $c$}{run itclust on the cells inside $c$\;}}
 \caption{itclust}
\end{algorithm}

This method is nice because the important low-dimensional representations \emph{within} a cluster may be quite different from the ones that help you distinguish \emph{between} two subclusters within that cluster.  This is what allows them to achieve the level of granularity that they have.

Unfortunately, it is a bit difficult to grasp the uncertainty involved in the process.  There is certainly noise in the data, and different practitioners of this basic procedure may have different understandings of what this noise is.  The result is that when Bosilijka Tasic (one of the lead experimentalists at Allen) gives two data scientists the same data, they each come back with different answers.  And they have no way to compare or evaluate these two different answers.  One essential question is ``what level of granularity does the signal-to-noise ratio support?''  The algorithm above, as stated, runs iteratively until subclusters cannot be distinguished via T-SNE.  But this is obviously a very subjective matter.  What to do?

We propose to help clarify some of these issues by building a fully generative model of gene expression, one that rigorously measures some kinds of uncertainty in every step of the process.  To be most useful to the Allen Institute, however, we also want this generative model to fit into the scientific paradigm under which they are operating.  In particular, the treatment of clusters should be hierarchical in nature, and it is desirable that the model have latent representations that roughly correspond to the low-dimensional representations found in the Allen algorithm.

\section{Model}

Towards this end, we propose the following generative model for gene expression of a single cell.  Let $H$ denote a tree of cell types.

\begin{enumerate}
\item $T \sim \Categorical(\pi)$, one of the leaves of $H$.  Let $t_0,t_1(T),t_2(T),\cdots ,t_\ell(T)$, indicate the nodes in the unique path in $H$ which leads from the root to the leaf $T$.  Here $t_0$ is the root of the tree, $\ell$ is the depth of the leaf $T$, and $t_\ell(T)=T$.   To consolidate notation, we take $T_i \triangleq t_i(T)$.
\item For each $i<\ell$ let $Z_{T_i} \sim \cN(\mu_{T_{i+1}},\Sigma_{T_{i+1}})$, and take $Z_{T_\ell} \sim \cN(0,I)$
\item For each gene $g$, let $\lambda_g \in \mathbb R^{p}$ denote a set of parameters.  These will be a deterministic function of the $Z$s:
\[
\lambda_g = f_g \left(\sum_{i=0}^{\ell} h_{T_i}Z_{T_i}\right)
\]
\item Conditioned on $Z$, the  $g$th gene of the cell be distributed according to some $p(x_g ; \lambda_g)$. 
\end{enumerate}

Let us unpack some of this:

\begin{itemize}
\item The latent representation $Z_i$ corresponds to features of that cell which are specific to all the cells which are descendents of $T_i$.
\item The the cell's subcluster $T_{i+1}$ effects the distribution of its latent features at the $i$th level, $Z_i$.  That is, the latent distribution on the $i$th latent representation is governed by what subcluster it is part of within $T_i$.
\item The distribution of the final latent representation, $Z_T$, is simply a standard normal.
\item The functions $h_{T_i}$ indicates how these cell-type-specific features alter the distribution on the gene expressions for cells in that branch of the tree.  
\end{itemize}

For inference, we use amortized variational inference.  In particular, we take the variational family that $T$ and all the $Z_{T_i}$ are independent, $T$ is some categorical, and $Z$s are isotropic gaussians.  We build a neural network to estimate the variational parameters from gene expressions.  Note that this network must estimate the parameters for a gaussian at \emph{each node} in the tree $H$.

\section{Empirical results}

At the current moment, we have only actually implemented this model in the case the $H$ is a tree with depth 1, i.e. the hierarchy is trivial.   We take the data model to be a zero-inflated negative binomial.  It looks fairly promising.  For a simple qualitative assessment, consider the following.

Scatterplot of the aggregate posterior distribution of $Z_{T_0}$, colored by posterior mean of the cluster assigment:

\includegraphics[width=0.4\textwidth]{pics/Z1}

Notice that there are only so many clusters that are really clearly distinct in this latent space.  However, let's focus on that cluster on the right-hand side.  What happens if we look at the $Z_{T_1}$ variables for those cells?  We get substantial subclustering apparent in the $Z_{T_2}$ space:

\includegraphics[width=0.4\textwidth]{pics/Z2}

\subsection{Held out log likelihood}

To quantitatively analyse our method, we used a held out log likelihood:

\begin{itemize}
\item Train the method on 75\% of the cells
\item For each of the remaining 25\% of the cells, select 75\% of the genes (``observed''):
  \begin{itemize}
  \item Compute the posterior mode of Z given the expression of those genes
  \item Compute the log probability of the observed genes given the posterior mode
  \item Compute the log probability of the unobserved genes given the posterior mode
  \end{itemize}
\item Average values over those 25\% of the cells
\end{itemize}

Here are the results for this ``Posterior mode predictive (PMP)'' approach:

\begin{tabular}{r|cc}
Method & PMP, observed & PMP, unobserved \\
Factor analysis $k=16$ & -4.2 & -4.2 \\
Factor analysis $k=64$ & -4.1 & -4.3 \\
HNBPF $k=2$ & -5.7 & -5.7 \\
HNBPF $k=16$ & -3.8 & -3.9 \\
HNBPF $k=64$ & -3.8 & -5.7 \\
ZINB-Vae $k=2$ & -3.6 & -3.6 \\
Smooth ZINB-Vae $k=2$ & -3.6 & -3.6 \\
Smooth Hierarchical ZINB $k=2+2$ & -3.5 & -3.5 \\
\hline
\end{tabular}

Here HNBPF is a negative-binomial-poisson-factorization approach, ZINB-Vae is based on a Michael Jordan paper, $k$ is the number of latent dimensions.  The results are certainly not mind-blowing, but its nice to see that we're at least not doing worse.

To do posterior inference on held out likelihood, our approach is to first set all of the ``unobserved'' entries to the average value for the corresponding genes, and apply our inference network unchanged to get estimates for the variational parameters.  We then refine these estimates with gradient descent.  We note that if we use completely random guesses to initialize the gradient descent, we often converge to inferior local minima; this shows the benefit of the amortized inference.

\subsection{Posterior predictive samples}

One limitation of the method as we have implemented it is that the posterior predictive distribution seems to be substantially underfitting.  To see what we mean, consider the following figure:

\includegraphics[width=0.8\textwidth]{pics/Z1gn}

Here we four rows and three columns of plots.  Each row corresponds to a different gene.  Each dot in each plot corresponds to a cell.  The position of the dot indicates the latent representation of the cell, as inferred by posterior inference.  The color indicates the amount of expression for the particular gene for that cell, measured three ways:

\begin{enumerate}
\item The first column shows the true expression of the appropriate gene for each cell (i.e the color of each dot is the gene expression for that cell)
\item The second column shows a posterior predictive sample of the gene expression for each cell
\item The third column shows the mean of such samples for each cell
\end{enumerate}

We see that the true expression doesn't look much like the posterior predictive expression for the Vip and Sst genes.  Something better is needed.  Possibly more layers.  Possibly more training.  Worth thinking about.

% \bibliographystyle{unsrt}
% \bibliography{refs}

\end{document}